### 2020-04-17 sections/discussion.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis, which assumes all arms in a given mode have the same pitch angle.

We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time. The assumptions of this model of spiral winding are also highly reductionist, and leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? This result is not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.

As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative. However, we believe that the methodlogy proposed here is a scalable, robust solution to the problems facing investigation of spiral morphology.

>>>>

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation, and better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.

We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence. \citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mas, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time. The assumptions of this model of spiral winding are highly reductionist, and it leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.

We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles within the limits specified above. We also assume that the galaxy sample is representative of the general spiral population, which is not necessarily the case. As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.

The methodlogy proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.

### 2020-04-17 sections/introduction.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

Arms of spiral galaxies are the source of the vast majority of star formation in the Universe, and spirals rearrange disc gas and can lead to the formation of disc-like bulges (e.g. \citealt{2004ARA&A..42..603K}). Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, \citealt{2015PhDT........14D}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure). These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.

@@

Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}. Recent research suggests that spirals are dynamical in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}. These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e. WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave. The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle and then dissapating to be reformed.

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation. \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time. They found that the pitch angles of their sample of 86 galaxies was consistent with this distribution, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

@@

Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations. Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.

>>>>

Arms of spiral galaxies are the source of the vast majority of star formation in the Universe, and spirals rearrange disc gas and can lead to the formation of disc-like bulges (e.g. \citealt{2004ARA&A..42..603K}). Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure). These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.

@@

Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}. Recent research suggests that spirals arms are transient in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}. These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e. WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave. The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissapating to be reformed.

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation. \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time. They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

@@

Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}). Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.

### 2020-04-17 sections/method.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the classifications of galaxies in this \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard. We remove any galaxies for which fewer than two spiral arms were identified, as these do not provide our model with any information about inter-arm scatter. This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.

@@

We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in this galaxy, $\phi_\mathrm{arm}$, are constants (giving logarithmic spirals) drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies. We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving

@@

The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo algorithm used (discussed later). The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution. Our likelihood function for $N$ arms, each of which with $n_\mathrm{arm}$ points is

>>>>

Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the classifications of galaxies in this \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard. We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.

@@

We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in the galaxy, $\phi_\mathrm{arm}$, are constant (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies. We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving

@@

The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later). The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution. Our likelihood function for $N$ arms, each of which with $n_\mathrm{arm}$ points is

### 2020-04-17 sections/results.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle. The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively. This is consistent with the standard error on the mean for a galaxy with $N$ arms,

@@

where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 1.0^\circ$. This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test \citep{10.2307/2286009} over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} and Kolmogorov-Smirnov tests in a similar manner for comparison.

\subsubsection{Pitch angle vs. Bulge size}
We define a bulge prominence from Galaxy Zoo 2 as Equation 3 in \citet{2019MNRAS.487.1808M}:

\begin{equation}
  B_\mathrm{avg} = 0.2p_\mathrm{just noticeable} + 0.8p_\mathrm{obvious} + 1.0p_\mathrm{dominant}.
\end{equation}

We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$. The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E(\phi_\mathrm{gal})$) and $B_\mathrm{avg}$ is -0.09. We see a slight correlation between $E(\phi_\mathrm{gal})$ and the fraction of volunteers who said a galaxy did not have a bulge (Pearson correlation coefficient of 0.19).

\comment{Do we need to do an Anderson-Darling test here?}

We separate our sample into ``weaker bulged galaxies'' and ``stronger bulged galaxies'' using , in order to examine whether they could be drawn from significantly different distributions. A marginalized two-sample Anderson-Darling test does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples. The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.

\subsubsection{Pitch angle vs. Bar Strength}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchical normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.

Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions (we reject the null hypothesis at the 1\% level for only 1\% of samples). The distribution of Anderson-Darling test statistics is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.

We do not account for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a non-axisymmetric focing.

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles for Bulge-dominated and Disc-dominated galaxies are drawn from the same distribution (top panel), and the results of the same test for strongly-barred vs unbarred galaxies (bottom panel).}
  \label{fig:ad-morphology-test}
\end{figure*}

@@

We aim to replicate their work here, using our sample and methods. We will make use of the marginalized tests described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.

@@

Testing the uniformity in cot of $\phi_\mathrm{gal}$, between the limits specified above, using a marginalized Anderson-Darling test, results in rejecting the null hypothesis at the 1\% level for 68\% of samples. This suggests that a cot uniform model is not a good fit to galaxy pitch angle, but is not conclusive enough for us to unilaterally reject winding governed by Equation \ref{eq:winding}. The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.

This inconclusive result is perhaps unsurprising; were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time. This is supported by the large spread in inter-arm pitch angles.

\subsubsection{Arm Pitch angle}
If we assume that spirals form and wind independantly in a galaxy and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits.

Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at the 10\% level for any of the possible realizations of arm pitch angle. The resulting distribution of Anderson-Darling statistics in the lower panel of Figure \ref{fig:ad-cot-test}. This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.

@@

This result is highly sensitive to the lower limit of $\phi_\mathrm{gal}$ chosen; decreasing it to 10\degree results in us rejecting the cot-uniform progenitor at greater than the 0.1\% level for every realization of the posterior. As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we should use the higher limits.

\comment{We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles. We also assume that the galaxy builder is representative of the general spiral population, which is not necessarily the case}

>>>>

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle. The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.9\degree}, which decreases to {6.8\degree} and {6.0\degree} for galaxies with three and four arms respectively. This is roughly consistent with the standard error on the mean for a galaxy with $N$ arms,

@@

where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$. This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.

\comment{what about for one-armed galaxies? (A: 9.86\degree)}

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test (\citealt{10.2307/2286009}, implemented in \textsc{Scipy}, \citealt{scipy-paper}) over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} test in a similar manner.

\subsubsection{Pitch angle vs. Bulge size}

Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}). Some studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e. \citealt{2017MNRAS.472.2263H}, \citealt{2019ApJ...873...85D}), while others have not found any significant correlation \citep{2019MNRAS.487.1808M}. We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:

\begin{equation}
  B_\mathrm{avg} = 0.2\times p_\mathrm{just\ noticeable} + 0.8\times p_\mathrm{obvious} + 1.0\times p_\mathrm{dominant},
\end{equation}

Where $p_\mathrm{just\ noticeable}$, $p_\mathrm{obvious}$ and $p_\mathrm{dominant}$ are the fractions of classifications indicating the galaxy's bulge was ``just noticeable'', ``obvious'' or ``dominant'' respectively.

We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$. The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00.

We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 83 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 54 galaxies), in order to test whether their pitch angles could be drawn from significantly different distributions. A marginalized two-sample Anderson-Darling test comparing the distributions of $\phi_\mathrm{gal}$ for the samples does not find evidence that galaxy pitch angles were drawn from different distributions: we reject the null hypothesis at the 1\% level for only 1\% of the samples. Similarly comparing arm pitch angles for galaxies in the different samples results in not rejecting the null hypothesis at the 1\% level for any of the samples. The distributions of Anderson-Darling test statistic for $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ are shown in the upper panel of Figure \ref{fig:ad-morphology-test} in blue and orange respectively.

These results may be caused by the galaxy sample not containing many galaxies with dominant bulges: $B_\mathrm{avg}$ only varied from 0.09 to 0.75 (the allowed maximum being 1.0), with only four galaxies having $B_\mathrm{avg} > 0.5$. The split point of 0.28 was also chosen to produce evenly sized comparison samples rather than from some physical motivation. However, the lack of any form of correlation strongly suggests that there is no evidence in our data for the link between bulge size and pitch angle observed in other studies.

\subsubsection{Pitch angle vs. Bar Strength}

One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}. In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which is widely viewed as a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.

We do not observe a correlation between \textit{pbar} and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05). Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($\mathrm{\textit{pbar}} < 0.2$), weakly-barred galaxies ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and strongly-barred galaxies ($\mathrm{\textit{pbar}} > 0.5$).

A marginalized k-sample Anderson-Darling tests on the distributions of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any of samples for either the test of $\phi_\mathrm{gal}$ or the test of $\phi_\mathrm{arm}$. The distributions of Anderson-Darling test statistic is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.

\comment{this paragraph is quite clunky, can we rephrase it?}

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results for strongly-barred vs unbarred galaxies (bottom panel).}
  \label{fig:ad-morphology-test}
\end{figure*}

We do not correct for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric focing. The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.

@@

We aim to replicate their work here, using our sample and methods. We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.

@@

Testing the uniformity of $\cot\phi_\mathrm{gal}$ between {15\degree} and {50\degree} using a marginalized Anderson-Darling test results in rejecting the null hypothesis at the 1\% level for just 5\% of samples, with a large spread in observed test values. The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.

This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements. This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples. As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we choose to keep the less strict limit of {15\degree}.

\subsubsection{Arm Pitch angle}
The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time. This is supported by the large observed spread in inter-arm pitch angles.

If we assume instead that spirals form and wind independantly in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits, rather than that of the galaxy's pitch angle as a whole.

Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at even the 5\% level for any of the possible realizations of arm pitch angle. The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}. This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.

@@
